---
title: "Hypothesis Testing"
author: "JamesQuacinella"
date: "April 19, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Z Scores

Here I will load a data set of heights and weights of people from Hong Kong:

```{r}
data <- read.csv('~/Desktop/height_data.csv')
head(data)
```

Lets inspect the height data:

```{r}
height <- data$height
hist(height) #histogram
```

Its easy in R to find the parameters for the population distribution:

```{r}
#population parameter calculations
N <- length(height)  # Population size

pop_sd <- sd(height) * sqrt( (length(height) - 1) / (length(height)) )   # Why this correction? sd in R calculates a
                                                                         # a sample sd, which has n-1 in the denom. 
                                                                         # Bessel's correction. Advanced

pop_mean <- mean(height) # Mean of the data

writeLines(paste("Population mean:", pop_mean))
writeLines(paste("Population   sd:", pop_sd))
```


Lets review z-scores. Z-scores allow us to talk about values in terms of how far that value is from the mean in units of standard deviation, What is the probability of getting a height of < 72inches? How do we convert that to a probability > 72in?

```{r}
data_point = 72

# z-score calculation
z <- (data_point - pop_mean) / pop_sd

p_less_than1 <- pnorm(data_point, pop_mean, pop_sd)    #using x, mu, and sigma
p_less_than2 <- pnorm(z)                               #using z-score of 2.107

# Convert to prob > the data point
p_greater1 <- 1 - p_less_than1
p_greater2 <- 1 - p_less_than2


# Probability of greater or equal to 72 is same no matter which way we calculate
writeLines(paste("Prob >= 72in:", p_greater1))
writeLines(paste("Prob >= 72in:", p_greater2))
```


## Hypothesis Testing

Here we are going to take a biased sample from the height data, focusing more on taller people than shorter. We do this to setup a hypothesis test to see if this sample is statistically different than the population.

```{r}
# Sample size
n <- 50

# tall-biased sample
cut <- 1:N
weights <- cut^.6
sorted_height <- sort(height)

set.seed(123)

height_sample_biased <- sample(sorted_height, size=n, prob=weights)
head(height_sample_biased)
```

The null hypothesis is that there is no difference in the means of the population and the sample. The alternative hypothesis, as opposed to the web page this was taken from, will be that the sample mean is greater than the population mean. 

Now that we have the sample, lets find the sample mean and sample sd:


```{r}
# Calculate the sample mean and sd
sample_mean <- mean(height_sample_biased)
sample_sd <- sd(height_sample_biased)

writeLines(paste("Sample mean:", sample_mean))
writeLines(paste("Sample   sd:", sample_sd))
```


```{r}
#t-stat
t <- (sample_mean - pop_mean) / (sample_sd / sqrt(n))
t
```

From this t score we can calculate the probability of getting a sample mean greater than or equal to what we observed:

```{r}
#p-value for t-test
p_value <- 1 - pt(t, n-1)
p_value
```

We can do this using t.test:

```{r}
# http://www.stat.columbia.edu/~martin/W2024/R2.pdf
t.test(height_sample_biased, alternative="greater", mu=pop_mean)
```