---
title: "PageRank"
author: "James Quacinella"
date: "April 19, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Transition Matrix

```{r cars}
n <- 6
A <- matrix(c(0,.5,.5,0,0,0,
              1/6,1/6,1/6,1/6,1/6,1/6,
              1/3,1/3,0,0,1/3,0,
              0,0,0,0,.5,.5,
              0,0,0,.5,0,.5,
              0,0,0,1,0,0), byrow=T, nrow=n)
A
```

Lets confirm the rows add to 1:

```{r}
rowSums(A)
```

Lets update our matrix with a decay matrix as per the notes:

```{r}
d <- 0.85

B <- d*A + ((1-d) / n)
B
```


Lets create a page rank vector, which represents probabilities:

```{r}
r <- rep(1/n, n) 
r
```

Lets do one iteration:

```{r}
r_new <- B %*% r
r_new
```


Wait ... why didn't r change?? If you think about how B is constructed with it rows summing to 1, then r via multiplication never changes! In order to see any changes in R, I had to swap the matrix multiplcaition order:


```{r}
r_new <- r %*% B
r_new
```

Lets do this repeatedly and find the page rank vector:


```{r}
library(expm) 
r_final <- r %*% ( B %^% 30 ) 
r_final
```


## Eigenvector Method

Lets try to solve this using eigenvectors:

```{r}
eigen(B)
```


Uh oh, the eigenvector with eigenvalue 1 doesn't look right at all! One thing to notice is that our method above is using matrix B in a way thats NOT an eigenvector problem. We need to do some math to switch the order of the terms:

$$ r = r * B $$
$$ r^T = (r * B)^T $$
$$ r^T = B^T * r^T $$

Hence, it seems like we need to look at eigenvectors of the transpose of this matrix:

```{r}
eigens <- eigen( t(B) )
eigens
```


```{r}
r_final_eigen <- eigens$vectors[,1]
r_final_eigen
```

Issue here is that this is not a pagerank vector since it does not represent probabilities that sum to 1. If r_final_eigen is a eigenvector solution to the equation, then so is any scalar multiple of it. We need to scale this eigen vector such that the components add to 1. NOTE: this is not 'normalizing' a vector

```{r}
r_final - r_final_eigen / sum(r_final_eigen)
```



```{r}
r_final_eigen <- r_final_eigen / sum(r_final_eigen)
r_final_eigen
```